\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2012-11-12 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
changed a few things in the manual
allows the use of multiple user-defined tomographic files instead of having the same and only "tomography_model.xyz" file, moves the TOMOGRAPHY_PATH flag in Par_file, organises the "done" items of the todo list

%%magnoni
For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
%
%For a tomographic model, the material definition format is:
%\begin{lyxcode}
%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
%\end{lyxcode}
%where
%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file. The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
%\texttt{src/generate\_databases/} directory.
%
\item [materials\_file] Contains the material associations for each element.
\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
\item [mesh\_file] Contains the mesh element connectivity.
\item [free\_surface\_file] Contains the free surface connectivity. You should put both the surface of acoustic regions and of elastic regions in that file;
that is, list all the element faces that constitute the surface of the model in that file.
\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
\end{description}

@@

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh  nparts  input\_directory output\_directory
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/}, and
     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition. These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
\end{itemize}
Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.

@@

@@

To implement your own external tomographic model, you must provide your own tomography file
\texttt{tomography\_model.xyz} and choose between two possible options:
(1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, or
(2) for more user control, set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{lyxcode}
\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
\end{lyxcode}
where
\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\texttt{material\_ID} a negative, unique identifier (i.e., -1),
\texttt{tomography} keyword for tomographic material definition,
\texttt{elastic} keyword for elastic material definition,
\texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.
The name of the tomography file is not used so far, rather change the filename (\texttt{TOMO\_FILENAME}) defined in the file \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases/} directory.

The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
The ASCII file \texttt{tomography\_model.xyz} that describes the tomography should be located in the \texttt{DATA/} directory.
The format of the file, as read from \texttt{model\_tomography.f90}, looks like Figure \ref{fig:tomography_file},
where
%
\begin{description}
\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
\end{description}
%
After the first four lines, in the file \texttt{tomography\_model.xyz} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
%fig
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
\par\end{centering}
\caption{Tomography file \texttt{tomography\_model.xyz} that describes an external Earth model and that is read by \texttt{model\_tomography.f90}. The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
\label{fig:tomography_file}
\end{figure}

>>>>

%%magnoni
For tomographic velocity models, please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} `Using external tomographic Earth models' for further details.
%
%For a tomographic model, the material definition format is:
%\begin{lyxcode}
%\texttt{domain\_ID} \texttt{material\_ID}  \texttt{tomography}  \texttt{name}
%\end{lyxcode}
%where
%\texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \texttt{material\_ID} a negative, unique identifier, \texttt{tomography} keyword for tomographic material definition and \texttt{name} the name of the tomography file. The name is not used so far, rather change the filename defined in the file \texttt{model\_tomography.f90} located in the
%\texttt{src/generate\_databases/} directory.
%
\item [materials\_file] Contains the material associations for each element. The format is: 
\begin{lyxcode}
\texttt{element\_ID} \texttt{material\_ID}
\end{lyxcode}
where \texttt{element\_ID} is the element identifier and \texttt{material\_ID} is a unique identifier, positive (for materials) or negative (for tomographic models).
\item [nodes\_coords\_file] Contains the point locations in Cartesian coordinates of the mesh element corners.
\item [mesh\_file] Contains the mesh element connectivity.
\item [free\_surface\_file] Contains the free surface connectivity. You should put both the surface of acoustic regions and of elastic regions in that file;
that is, list all the element faces that constitute the surface of the model in that file.
\item [absorbing\_surface\_file\_xmax] Contains the surface connectivity of the absorbing boundary surface at the Xmax.
\item [absorbing\_surface\_file\_xmin] Contains the surface connectivity of the absorbing boundary surface at the Xmin.
\item [absorbing\_surface\_file\_ymax] Contains the surface connectivity of the absorbing boundary surface at the Ymax.
\item [absorbing\_surface\_file\_ymin] Contains the surface connectivity of the absorbing boundary surface at the Ymin.
\item [absorbing\_surface\_file\_bottom] Contains the surface connectivity of the absorbing boundary surface at the bottom.
\end{description}

@@

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./xdecompose\_mesh  nparts  input\_directory output\_directory
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/}, and
     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition. These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
\end{itemize}
Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.

@@

\item [{\texttt{TOMOGRAPHY\_PATH}}] Directory in which the tomography files are stored for using external 
tomographic Earth models (please read Chapter \ref{cha:-Changing-the} and Section \ref{sec:Using-tomographic} 
`Using external tomographic Earth models' for further details.).

@@

To implement your own external tomographic model(s), you must provide your own external tomography file(s), 
and choose between two possible options:\\
\indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = tomo}, \\ 
or for more user control: \\
\indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL = default}, define the negative \texttt{material\_ID} identifier for each element in the file \texttt{MESH/materials\_file} and use the following format in the file \texttt{MESH/nummaterial\_velocity\_file} when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{lyxcode}
\texttt{domain\_ID} \texttt{material\_ID} \texttt{tomography} \texttt{elastic} \texttt{file\_name} 1
\end{lyxcode}
where: \\
\indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials, \\
\indent \texttt{material\_ID} a negative, unique identifier (i.e., -1,-2,...), \\
\indent \texttt{tomography} keyword for tomographic material definition, \\
\indent \texttt{elastic} keyword for elastic material definition, \\
\indent \texttt{file\_name} the name of the tomography file and 1 a positive unique identifier.\\

The external tomographic model is represented by a grid of points with assigned material properties and homogeneous resolution along each spatial direction x, y and z.
The ASCII file \texttt{file\_name} that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH} directory, set in the \texttt{Par\_file}.
The format of the file, as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases} directory, looks like Figure \ref{fig:tomography_file},
where
%
\begin{description}
\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates of the initial and final tomographic grid points along the x direction (in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates of the initial and final tomographic grid points along the y direction (in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates of the initial and final tomographic grid points along the z direction (in the mesh units, e.g., $m$);
\item [\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}] the spacing between the tomographic grid points along the x, y and z directions, respectively (in the mesh units, e.g., $m$);
\item [\texttt{NX}, \texttt{NY}, \texttt{NZ}] the number of grid points along the spatial directions x, y and z, respectively; \texttt{NX} is given by [(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}]+1; \texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the y and z directions, respectively;
\item [\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}] the minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs} (in $m \, s^{-1}$) and of the density \texttt{rho} (in $kg \, m^{-3}$); these values could be the actual limits of the tomographic parameters in the grid or the minimum and maximum values to which we force the cut of velocity and density in the model.
\end{description}
%
After the first four lines, in the file \texttt{file\_name} the tomographic grid points are listed with the corresponding values of \texttt{vp}, \texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate (from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X}) for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to \texttt{END\_Z}, with step of \texttt{SPACING\_Z}).
%fig
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=1.0\textwidth]{figures/tomo_file.jpg}
\par\end{centering}
\caption{Tomography file \texttt{file\_name} that describes an external Earth model. The coordinates x, y and z of the grid, their limits ORIG\_X, ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X, SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh (e.g., UTM coordinates in meters).}
\label{fig:tomography_file}
\end{figure}
